---
title: "Spike Computational cost"
author: "Gabriel Ristow Cidral / Sara Marín-Lopez"
date: " 13 / 02 / 20"
output: html_document
     code_folding: hide
---

This spike is intended to provide ideas how of to make your code more efficient. 

It will approach 3 methods: Sma rt samples, parallel processing, modeling without caret, opmitization of random forest (mtry)


```{r part 1, message=F}

pacman::p_load(readr, dplyr, caret, plotly)

setwd("C:/Users/gabri/Desktop/Ubiqum/R/Deep_Analytics_and_Visualization/Task_3/Wifi_fingerprinting")

# setwd(
# "C:/Users/gabri/Desktop/Ubiqum/R/Deep_Analytics_and_Visualization/Task_3/"
# )
train <- read_csv("trainingData.csv", na = c("N/A"))

sample <- train %>% group_by(FLOOR, BUILDINGID) %>% sample_n(10)
table(sample$BUILDINGID)
table(sample$FLOOR)


```


```{r plotly, echo=F}

p <- sample %>%
plot_ly(type = "scatter3d",
        x = ~ LATITUDE,
        y = ~ LONGITUDE,
        z = ~ FLOOR,
        text = ~ sample,
        hoverinfo = 'text',
        mode = 'markers',
        transforms = list(
          list(
            type = 'filter',
            target = ~ BUILDINGID,
            operation = '=',
            value = unique(sample$BUILDINGID)[1]
            )
          )
        ) %>% layout(updatemenus = list(list(
          type = 'dropdown', active = 0, 
          buttons = list(list(
            method = "restyle",
            args = list("transforms[0].value", unique(sample$BUILDINGID)[1]),
            label = unique(sample$BUILDINGID)[1]),
            list(
              method = "restyle",
              args = list("transforms[0].value", unique(sample$BUILDINGID)[2]),
              label = unique(sample$BUILDINGID)[2]),
            list(method = "restyle",
                 args = list("transforms[0].value", unique(sample$BUILDINGID)[3]),
                 label = unique(sample$BUILDINGID)[3])
            )
          )))



table(sample$BUILDINGID)



p

```

## PARALLEL PROCESSING ##
A computer usually has multiple cores. Tipically, R is going to use only one of them, but we can increase this number, allowing us to execute more computations at the same time. 

**How to do it on Windows**
<ul>
  <li>Install the doParallel package </li>
  <li>Check how many cores you have with the function **<span style="color:MIDNIGHTBLUE">detectCores()</span>**.</li>
  <li>Save the number of cores that you would like to execute with the function **<span style="color:MIDNIGHTBLUE ">makeCluster()</span>**. A good practice is to leave one for other tasks. </li>
  <li>Register the cluster with the function **<span style="color:MIDNIGHTBLUE">registerDoParallel()</span>**</li>
</ul>

**How to do it on Mac/Linux**
<ul>
  <li>Install the doMC package </li>
  <li>Check how many cores you have with the function **<span style="color:MIDNIGHTBLUE">getDoParWorkers()</span>** </li>
  <li>Save the number of cores that you would like to execute with the function **<span style="color:MIDNIGHTBLUE">makeCluster()</span>**. A good practice is to leave one for other tasks. </li>
  <li>Register the cluster with the function **<span style="color:MIDNIGHTBLUE">registerDoMC()</span>**</li>
</ul>

Now you can apply parallel processing! For example, you can use it in the cross validation or in the RF with the parameter "allowParallel = TRUE". 

**Challenge: Train the same sample with parallel processing**
```{r eval=FALSE,message=FALSE}
# Load the library
library(doParallel)

# Check number of cores
detectCores()

# Save the number of cores I'm going to use
cluster <- makeCluster(detectCores() - 1)

# Register the cluster
registerDoParallel(cluster)

# Apply it on the cross validation
fitControl <- trainControl(method = "repeatedcv", number = 10, repeats = 3, allowParallel = TRUE)

```

## SPECIFIC PACKAGES ##
### Random Forest: package randomForest ###
This is the most usual package for training a random forest. It's very user friendly and robust. If you want to learn more about other packages check this [resource](https://www.linkedin.com/pulse/different-random-forest-packages-r-madhur-modi/).

Let's see which are the main parameters of the function **<span style="color:MIDNIGHTBLUE">randomForest</span>**:
<ul>
  <li> ntree: number of trees to grow </li>
  <li> mtry: how many random variables will be selected to grow in a single tree </li>
  <li> importance: should importance of predictors be assessed? *Keep in mind that if your data includes categorical variables with different number of levels, random forests are biased in favor of those variables with more levels.* </li>
</ul>

Another useful function from this package is **<span style="color:MIDNIGHTBLUE">tuneRF()</span>**. Starting with the default value of mtry, it searchs for the optimal value. 

**Your turn! Try to obtain the best mtry for your data and train a random forest using this package.**

```{r eval=FALSE, message=FALSE }
# Load package
library(randomForest)

# Get the best mtry
bestmtry_rf<-tuneRF(training[WAPS], training$LONGITUDE, ntreeTry=100,stepFactor=2,improve=0.05,trace=TRUE, plot=T) 

# Saving the waps in a vector
WAPs<-grep("WAP", names(train), value=T))

# Train a random forest using that mtry
system.time(longitude_rf<-randomForest(y=train$LONGITUDE,x=train[WAPs],importance=T,method="rf", ntree=100, mtry=104))
```

### KNN: kknn package ###
This is the most usual package for training a knn. Let's see which are the main parameters of the function **<span style="color:MIDNIGHTBLUE">train.kknn</span>**:
<ul>
  <li> ks: number of neighbours </li>
</ul>

**Your turn! Try to train a knn with this package**
```{r eval=FALSE, message=FALSE}
# Load the package
library(knnn)

# Train the model
system.time(longitude_knn<-kknn())

```

### SVM:  ###


## SAVING AND LOADING MODELS ##
You can save your best models to a file. This way, you will be able to load/share them later. 
<ul>
<li> For saving a model you can mainly use two functions: **<span style="color:MIDNIGHTBLUE">save(___.rda)</span>** or **<span style="color:MIDNIGHTBLUE">saveRDS(____.rds)</span>** </li>
<li> For loading a model you will need to use **<span style="color:MIDNIGHTBLUE">load(____.rda)</span>** or **<span style="color:MIDNIGHTBLUE">readRDS(____.rds)</span>**
</ul>

**Your turn! Try to save and load some models.**
```{r eval=FALSE, message=FALSE}
# Save a model
saveRDS(RF_Model, file="RF_Model.rds")

# Load a model
final_model<-readRD("RF_Model.rds")
```

### COMPARE RUNNING TIMES ###
Let's compare how takes a random forest with caret, the function randomForest and parallel processing

You will need to use the function **<span style="color:MIDNIGHTBLUE">microbenchmark()</span>** from the package benchmark. 

*Example: results<-microbenchmark("model"= {"lm=formula"},*

** Do it by yourself **
```{r eval=FALSE, message=FALSE}
```


